---
title: "Scholar WordGraph"
author: "Tommi Suvitaival, Steno Diabetes Center Copenhagen, tommi.raimo.leo.suvitaival@regionh.dk"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    fig_width: 8
    fig_height: 8
    dev: jpeg
    toc: yes
  github_document:
    fig_width: 8
    fig_height: 8
    dev: jpeg
    toc: yes
always_allow_html: true
---

# Introduction

This is a tutorial script to producing an interactive visualization of words in a publication list extracted from [Google Scholar](https://scholar.google.com/). The appearance of the words is dependent on the frequency of their occurrence, and on which words they co-occur with in the titles of the listed articles.

The code is for R and it is, among others, based on packages [ggplot2](https://ggplot2.tidyverse.org/), [plotly](https://plotly.com/r/) and [htmlWidgets](https://www.htmlwidgets.org/).

This document is part of the Github repository [ScholarWordGraph](https://github.com/tommi-s/ScholarWordGraph) by [Tommi Suvitaival](https://tommi-s.com/).

<iframe 
  id="igraph" 
  scrolling="no" 
  style="border:none;" 
  seamless="seamless" 
  src="output-dark/index.html" 
  height="800" 
  width="800"
>
widget
</iframe>

# Preparations

1. Open a Google Scholar profile with browser (e.g., https://scholar.google.com/citations?user=SGgP0VQAAAAJ ).
1. Go to the bottom of the page and click "SHOW MORE" to show all entries on the page.
1. Go again to the bottom of the newly-expanded page.
1. Select everything in the table, starting from the year of the last entry (on the bottom-right corner) and ending at the title of the first entry (on the top-left corner).
1. Copy the selected text.
1. Paste the text to a text editor and save as "data/publications.txt" (see [example](https://github.com/tommi-s/ScholarWordGraph/blob/main/data/publications.txt) in the present repository).

# Load the Data

```{r Load}

data.loaded <- 
  readr::read_delim(
    file = "data/publications.txt",
    delim = "\t",
    escape_double = FALSE,
    col_names = FALSE,
    trim_ws = TRUE
  )

```

# Format the Data

## Create a Data Frame of Articles

* Create a matrix with articles as rows and title, authors and journal as the three columns.
* Convert to a data frame.

```{r Matrix}

data <- unlist( data.loaded )

data <-
  matrix(
    data = unlist( data ),
    ncol = 3,
    byrow = TRUE
  )

colnames( data ) <-
  c(
    "Title",
    "Authors",
    "Journal"
  )

data <-
  data.frame(
    data,
    stringsAsFactors = FALSE
  )

```

## Convert Titles to Lower Case

* To match identical words with different case, convert all to lowercase.

```{r Tolower}

data$"Title.lower" <-
  tolower( data$"Title" )

```

## Define Multi-Word Terms

* Define word sequences that should be considered as a single term by adding replacing whitespace with underscore.
* E.g., "type 1 diabetes" is one term.

```{r Multi-Word-Terms}
  
data$"Title.lower" <-
  stringr::str_replace_all(
    string = data$"Title.lower",
    # pattern = "type 1 diabete",
    pattern = "type 1 diabetes",
    replacement = "type_1_diabetes"
  )

data$"Title.lower" <-
  stringr::str_replace_all(
    string = data$"Title.lower",
    pattern = "type 2 diabetes",
    replacement = "type_2_diabetes"
  )

```

## Omit Special Characters

* Omit punctuation and special characters from the titles.

```{r Omit-Punctuation}

words <-
  stringr::str_split(
    string = data[ , "Title.lower" ],
    pattern = "\\s"
  )

words <- 
  lapply(
    X = words,
    FUN = tolower
  )

words <-
  lapply(
    X = words,
    FUN = stringr::str_replace_all,
    pattern = "(\\()|(\\))|(\\:)|(\\,)|(\\.)",
    replacement = ""
  )

table.words <- 
  sort(
    x = table( unlist( words ) ),
    decreasing = TRUE
  )

head( table.words )

```

## Extract Unique Words

* List all unique words that appear in the titles.

```{r Words-Unique}

words.unique <-
  sort(
    unique(
      unlist( words ) )
  )
  
```

## Define List of Blocked Words

* List general words that are not of interest.
* Add plain numbers to this list.

```{r Blocklist}

blocklist <-
  c(
    "and",
    "an",
    "after",
    "are",
    "as",
    "at",
    "based",
    "by",
    "do",
    "during",
    "for",
    "from",
    "in",
    "is",
    "of",
    "on",
    "not",
    "the",
    "to",
    "through",
    "with",
    "without",
    "...",
    letters
  )

blocklist <- 
  c(
    blocklist,
    words.unique[ grepl( x = words.unique, pattern = "^[0-9]+$" ) ]
  )

```

## Omit Blocked Words

* Omit the blocked words from the words list of each article.

```{r Omit-Blocklisted}

words <-
  lapply(
    X = words,
    FUN = function( x ) {
      x[ !( x %in% blocklist ) ]
    }
  )

table.words <- 
  sort(
    x = table( unlist( words ) ),
    decreasing = TRUE
  )

head( table.words )

```

# Re-Extract Unique Words

```{r Words-Unique-Update}

words.unique <-
  sort(
    unique(
      unlist( words ) )
  )

```

# Create the Words-by-Articles Occurrence Matrix

* Matrix that has value 1 if the word (row) appears in the article (column) and 0 otherwise.

```{r Words-by-Articles}

wba <-
  array(
    data = 0,
    dim = c( length( words.unique ), nrow( data ) )
  )

rownames( wba ) <- words.unique
colnames( wba ) <- data[ , "Title" ]

for ( i in 1:nrow( data ) ) {
  
  wba[ words[[ i ]], i ] <- 1
  
}

```

## Log-Transform the Occurrences

* Trasform the occurrences with log (1+x) for a more Gaussian-like distribution.

```{r Log-Transform}

wba.log1px <- log10( 1 + wba )

```

## Autoscale the Occurrence Profiles

* Scale the data for PCA.

```{r Autoscale}

wba.norm <- scale( wba.log1px )

```

# Principal Component Analysis

## Compute PCA

```{r PCA-Compute}

result.pca <-
  princomp(
    x = wba.norm
  )

```

## Plot PCA

```{r PCA-Plot}

library( "ggfortify" )

plot <-
  autoplot( 
    result.pca
    # ,
    # loadings = TRUE,
    # loadings.label = TRUE
  )

print( plot )

```

# Cluster the Words

* Using a fixed number of eight clusters for a suitable visual representation.

```{r Clustering}

result.clustering <-
  cluster::pam(
    x = wba.norm,
    k = 8
  )

```

# Figure

## Prepare Results for the Visualization

* Extract PCA-scores of the two first components.
* Extract cluster assignments for each data point.
* Compute each data point's Euclidean distance from the origin.
* Wrap multi-word terms to multiple lines for better visual impression.
* Create a text string listing all the article titles, where the term appears.

```{r Data-for-Figure}

data.plot <-
  data.frame(
    Word = rownames( result.pca$"scores" ),
    result.pca$"scores"
  )

data.plot$"Cluster" <- 
  factor(
    x = result.clustering$"clustering",
    levels = 1:length( unique( result.clustering$"clustering" ) ),
    labels = 
      c(
        "Cluster (C) 1",
        paste( "C", 2:length( unique( result.clustering$"clustering" ) ) )
      )
  )
  
data.plot$"Distance" <- 
  sqrt( data.plot$"Comp.1"^2 + data.plot$"Comp.2"^2 ) + 1

data.plot$"Term" <-
  stringr::str_replace_all(
    string = data.plot$"Word",
    pattern = "\\-",
    replacement = "-\n"
  )

data.plot$"Term" <-
  stringr::str_replace_all(
    string = data.plot$"Word",
    pattern = "\\_",
    replacement = "\n"
  )

data.plot$"Publications" <-
  apply(
    X = wba[ data.plot$"Word", ] == 1,
    MAR = 1,
    FUN = function( x ) {
      paste(
        "\n",
        names(
          which( x )
        ),
        collapse = "\n"
      )
    }
  )

```

## Define the Figure

* Visualization with ggplot2.
* PCA components 1 and 2, respectively, on the y- and x-axes.
* Term name as text label.
* Distance from origin as (increasing) size.
* Cluster assignment as color.
* Add jitter to reduce overlap of similar terms.
* Use pseudo-log-valued axes for reduced outliers.

```{r Figure-Define}

jitter <- 1

palette <- RColorBrewer::brewer.pal( n = 9, "Set1" )[ -6 ]
palette <- palette[ 1:nlevels( data.plot$"Cluster" ) ]

plot <-
  ggplot2::ggplot(
    data = data.plot,
    mapping =
      ggplot2::aes(
        x = Comp.2,
        y = Comp.1,
        label = Term,
        size = Distance,
        text = Publications
        ,
        color = Cluster,
      )
  ) +
  ggplot2::geom_text(
    position =
      ggplot2::position_jitter(
        height = jitter,
        width = jitter
      )
  ) +
  ggplot2::scale_x_continuous( trans = "pseudo_log" ) +
  ggplot2::scale_y_continuous( trans = "pseudo_log" ) +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank()
  ) +
  ggplot2::xlab( label = "\n\n\nMap of my research topics" ) +
  ggplot2::ylab( label = "" ) +
  ggplot2::scale_color_manual( values = palette )

```

## Show the (Passive) Figure

```{r Figure-Show}

plot

```

## Create the Interactive Figure

* Make ggplot2 figure interactive with the **ggplotly**-function from the **plotly**-package.
* Show the term and the articles, where it appears, as text (tooltip) when hovering over a data point.
* Use the layout-function for fine-tuning of the legend.

```{r Figure-Interactive-Define}

library( tidyr )

plot.interactive <-
  plotly::ggplotly( 
    p = plot,
    tooltip = c( "label", "Publications" )
  ) %>%
  plotly::layout(
    legend = 
      list(
        orientation = "h",
        title = list( text = "Cluster<br>" ),
        x = quantile( x = data.plot$"Comp.2", probs = 0.67 ),
        y = min( data.plot$"Comp.1" )
      )
  )

```

## View the Interactive Figure

```{r Figure-Interactive-Show}

plot.interactive

```

## Export HTML Widget

* Create a stand-alone html-widget of the figure with the **partial_bundle** function from the **plotly**-package.
* Save the widget with the **saveWidget**-function from the **htmlWidgets**-package.

```{r Figure-Interactive-Export}

widget <- plotly::partial_bundle( plot.interactive )

# dir.create( "output" )

# htmlwidgets::saveWidget( widget, "output/index.html" )

```

# View the Exported Widget

* View the exported (saved) stand-alone widget with in an **iframe** in a html-document.

<iframe 
  id="igraph" 
  scrolling="no" 
  style="border:none;" 
  seamless="seamless" 
  src="output/index.html" 
  height="800" 
  width="800"
>
widget
</iframe>

# Figure with a Dark Layout

## Define the Figure

* Create the same figure with a dark layout, suitable for dark webpages.

```{r Figure-Dark-Define}

plot.dark <-
  plot +
  ggplot2::theme(
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank(),
    axis.title = ggplot2::element_text( color = "gray" ),
    legend.background = ggplot2::element_rect( color = NA, fill = "black" ),
    legend.key = ggplot2::element_rect( color = "gray",  fill = "gray" ),      
    legend.text = ggplot2::element_text( color = "gray" ),
    legend.title = ggplot2::element_text( color = "gray" ),
    panel.background = ggplot2::element_rect( fill = "black", color  = NA ),
    panel.grid = ggplot2::element_blank(),
    plot.background = ggplot2::element_rect( color = "black", fill = "black" )
  )

```

## Show the (Passive) Figure

```{r Figure-Dark-Show}

plot.dark

```

## Create the Interactive Figure

```{r Figure-Interactive-Dark-Define}

library( tidyr )

plot.dark.interactive <-
  plotly::ggplotly( 
    p = plot.dark,
    tooltip = c( "label", "Publications" )
  ) %>%
  plotly::layout(
    legend = 
      list(
        orientation = "h",
        title = list( text = "Cluster<br>" ),
        x = quantile( x = data.plot$"Comp.2", probs = 0.67 ),
        y = min( data.plot$"Comp.1" )
      )
  )

```

## View the Interactive Figure

```{r Figure-Interactive-Dark-Show}

plot.dark.interactive

```

## Export HTLM Widget

```{r Figure-Interactive-Dark-Export}

widget <- plotly::partial_bundle( plot.dark.interactive )

# dir.create( "output-dark" )

# htmlwidgets::saveWidget( widget, "output-dark/index.html" )

```

## View the Exported Widget

<iframe 
  id="igraph" 
  scrolling="no" 
  style="border:none;" 
  seamless="seamless" 
  src="output-dark/index.html" 
  height="800" 
  width="800"
>
widget
</iframe>

# SessionInfo

```{r SessionInfo}

utils::sessionInfo()

```
